The Evolution of Phenotypically Invariant Genetic Networks

### 2017-02-09 kalman_walk.tex
rewrote a fitness function.

\documentclass[11 pt]{article}
\usepackage{amsmath, amssymb, color, xcolor}
\usepackage{graphicx, wrapfig, float, caption}
\usepackage{fullpage}
\usepackage[backref=page, hidelinks, colorlinks=true, citecolor=blue!60!black!100]{hyperref}
\usepackage{tikz}
\usetikzlibrary{arrows.meta, shapes}
\usepackage{caption, subcaption}
\usepackage{natbib} % gives us \citet: Author (year) and \citep: (Author; year)
\usepackage{authblk}

\newcommand{\plr}[1]{{\color{blue}\it #1}}
\newcommand{\jss}[1]{{\color{olive}\it #1}}
% \newcommand{\ddt}{\frac{d}{dt}}
\newcommand{\ddt}{\dot}
\newcommand{\ro}{{ro}}
\newcommand{\nro}{{\bar{r}o}}
\newcommand{\rno}{{r\bar{o}}}
\newcommand{\nrno}{{\bar{r}\bar{o}}}
\newcommand{\reachable}{\mathcal{R}}
\newcommand{\unobservable}{\bar{\mathcal{O}}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\pda}{\frac{\partial}{\partial A_{ij}}}

@@

    \item Consider options for fitness function: Euclidean? Sign of eigenvalues?  Time domain or frequency?
        First try on simple concrete examples.

@@

\section{Including Deleterious Mutations and a Fitness Function}

@@

\begin{align}\label{eq:fitness}
  \Phi(\bar{\Sigma}) := \left\{ \begin{array}{ll}   e^{- | H(0) - \tilde{H}(0) | = - \left| \frac{k}{\lambda_{1} \lambda_{2}} - \frac{\tilde{k}}{\tilde{\lambda}_{1} \tilde{\lambda}_{2}} \right|}    &\text{if } \lambda_{1,2} < 0 \text{ and } \in \mathbb{R}, \ \det(A) \neq 0 \\
  0   &\text{if } \lambda_{1,2} \geq 0 \ \text{ or } \ \notin \mathbb{R}, \ \det(A) = 0 \end{array} \right.
\end{align}
and let $\upsilon$ be the magnitude and direction (sign) of a mutation on regulatory interaction strength. The fitness function (\ref{eq:fitness}) may me appropriate if mutations only slightly alter the dynamics of the system, as in Figure \ref{fig:mut}.

Further let $\mu(\Sigma, A_{ij})$ be resulting gene regulatory network following a mutation of matrix element $A_{ij}$ to system $\Sigma$ (which contains matrix $A$).

@@

>>>>

\documentclass[11 pt]{article}
\usepackage{amsmath, amssymb, color, xcolor}
\usepackage{graphicx, wrapfig, float, caption, dsfont}
\usepackage{fullpage}
\usepackage[backref=page, hidelinks, colorlinks=true, citecolor=blue!60!black!100]{hyperref}
\usepackage{tikz}
\usetikzlibrary{arrows.meta, shapes}
\usepackage{caption, subcaption}
\usepackage{natbib} % gives us \citet: Author (year) and \citep: (Author; year)
\usepackage{authblk}

\newcommand{\plr}[1]{{\color{blue}\it #1}}
\newcommand{\jss}[1]{{\color{olive}\it #1}}
% \newcommand{\ddt}{\frac{d}{dt}}
\newcommand{\ddt}{\dot}
\newcommand{\ro}{{ro}}
\newcommand{\nro}{{\bar{r}o}}
\newcommand{\rno}{{r\bar{o}}}
\newcommand{\nrno}{{\bar{r}\bar{o}}}
\newcommand{\reachable}{\mathcal{R}}
\newcommand{\unobservable}{\bar{\mathcal{O}}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\pda}{\frac{\partial}{\partial A_{ij}}}
\newcommand{\ind}{\mathds{1}}

@@

    \item Consider options for fitness function: Euclidean? Sign of eigenvalues?  Time domain or frequency?
      First try on simple concrete examples. \jss{see \ref{sec:del}}

@@

\section{Including Deleterious Mutations and a Fitness Function}\label{sec:del}

@@

\begin{align}\label{eq:fitness}
  \Phi(\bar{\Sigma}) := \left\{ \begin{array}{ll}   e^{- \left| H(0) - \tilde{H}(0) \right| = - \left| \frac{k}{\lambda_{1} \lambda_{2}} - \frac{\tilde{k}}{\tilde{\lambda}_{1} \tilde{\lambda}_{2}} \right|}    &\text{if } \lambda_{1,2} < 0 \text{ and } \in \mathbb{R}, \ \det(A) \neq 0 \\
  0   &\text{if } \lambda_{1,2} \geq 0 \ \text{ or } \ \notin \mathbb{R}, \ \det(A) = 0 \end{array} \right.
\end{align}
and let $\upsilon$ be the magnitude and direction (sign) of a mutation on regulatory interaction strength. The fitness function (\ref{eq:fitness}) may me appropriate if mutations only slightly alter the dynamics of the system, as in Figure \ref{fig:mut}.

Further let $\mu(\Sigma, A_{ij})$ be the resulting gene regulatory network following a mutation of matrix element $A_{ij}$ to system $\Sigma$ (which contains matrix $A$).

@@

 The transfer function at $z=0$ for any GRN with mutations is, 
  \begin{align*}
    H(0, \begin{bmatrix} A_{11} + \upsilon & A_{12} + \varepsilon \\ A_{21} + \delta & A_{22} + \omega \end{bmatrix}) = \frac{A_{22} - A_{12} + \omega - \varepsilon}{ \det(A) + \omega A_{11} - \delta A_{12} - \varepsilon A_{21} + \upsilon A_{22} + \omega \upsilon - \varepsilon \delta }
  \end{align*}

%  \begin{align*}
%    \frac{\left(A_{22}-A_{12}\right) + \left( \displaystyle\sum_{\jmath = 1}^{\ind_{\omega}} \omega_{\jmath} - \displaystyle\sum_{\jmath = 1}^{\ind_{\varepsilon}} \varepsilon_{\jmath} \right)}{\displaystyle\det(A) + \displaystyle\sum_{\jmath=1}^{\ind_{\omega}} \omega_{\jmath} A_{11} - \sum_{\jmath=1}^{\ind_{\delta}} \delta_{\jmath} A_{12} -  \sum_{\jmath=1}^{\ind_{\varepsilon}} \varepsilon_{\jmath} A_{21} + \sum_{\jmath=1}^{\ind_{\upsilon}} \upsilon_{\jmath} A_{22} + \sum_{\jmath=1}^{\ind_{\omega}} \omega_{\jmath} \sum_{\jmath=1}^{\ind_{\upsilon}} \upsilon_{\jmath} - \sum_{\jmath=1}^{\ind_{\varepsilon}} \varepsilon_{\jmath} \sum_{\jmath=1}^{\ind_{\delta}} \delta_{\jmath} }
%  \end{align*}

  If we want to know how new mutations fix in the population, we can explicitly show all mutational paths that could compensate for fixed mutations at sites $A_{ij}$. In a population, not at a phenotypic optimum, we are interested in the combination of mutations with varying magnitudes and locations that are most likely to fix.

  \begin{align*}
    \frac{\left(A_{22}-A_{12}\right) + \left( \displaystyle\sum_{\jmath = 1}^{\ind_{\omega}} \omega_{\jmath} - \displaystyle\sum_{\jmath = 1}^{\ind_{\varepsilon}} \varepsilon_{\jmath} \right)}{\det(A) + \displaystyle\sum_{\mu=\{\omega , \delta , \varepsilon , \upsilon \}}^{\ind_{\mu}} \mu_{\jmath} (-1)^{i+j} A_{ij} + \sum_{\jmath=1}^{\ind_{\omega}} \omega_{\jmath} \sum_{\jmath=1}^{\ind_{\upsilon}} \upsilon_{\jmath} - \sum_{\jmath=1}^{\ind_{\varepsilon}} \varepsilon_{\jmath} \sum_{\jmath=1}^{\ind_{\delta}} \delta_{\jmath} }
  \end{align*}
